HiC Contact Map Comaprison Using Graphlet Approach

### 2018-04-28 papers/manuscripts/manuscript.tex
modified papaer

\section{results and discussions}

\begin{figure*}
    \centering
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{figures/orbit-distances_chr14.png}
    \end{subfigure}
    \caption{   
        \textbf{Pair-wise graphlet signature difference for
        all 276 contact maps:}
        Each point on a graph is the result of averaging all
        the distances across all loci of that contact map.
        ($\bar{\mathbf{d}}^{\scriptscriptstyle A,B}_{i,j} \quad
        \forall i,j \in \{1 ... 23\} \quad \& \quad j \ge i$).  
     }
    \label{fig:results_all}
\end{figure*}

@@

As an alternative, we can compare graphlets by calculating
how often certain graphlets occur in them. By doing so
we measure the frequency distribution
of certain spatial structures in
each contact map. As a result of this approach
we can compare contact maps by calculating the 
correlation between their orbit distributions.
A higher correlation would mean higher similarity
in terms of spatial structure between the loci
involved.

We caclulated pair-wise MIC values 
for each orbit in each of the 276
contact maps from MIT, ALL, RL, and CALL4 data separately. 
As an example,
figure \ref{fig:contact_maps_correlations_chr14}
shows average orbit correlations across contact maps
for chromosome 14.
It corroborates the
concolusion we previously made that ALL and RL cell lines 
are more similar to each other than to other pairs by showing
significanly larger (refer to supplementary material for statistical tests)
orbit correlations between contact maps.
It is worth mentioning that interchromosomal thresholded contact maps 
represent
a bipartide graph with the loci from each chromosome on one side. Due to this
bipartide nature of the graphs in inter-chromosomal maps,
count of certain orbits is always 0, resulting in
a correlation values of 0 for them as well.
We ignored these values  when we calculated averages across orbits 
in figure \ref{fig:contact_maps_correlations_chr14} since they
would result in a bias towards zero in averages. You can see the bias in 
figure \ref{fig:orbits_correlations_all} where average correlations of orbits
$\mathbb{Q} = \{3, 9, 10{\text -}14, 20{\text -}34, 39{\text -}48, 51{\text -}72\}$ 
are close to zero. In fact all correlations
corresponding to these orbits are 0 except for the ones between the same 
chromosomes.

This approach also allows for comparison of orbits
themsleves. By comparing average oribit correlations
across all contact maps of two cell lines, we can
have a measure of how similar certain orbits are
between cell lines. This can give us a 
\textit{global} measure
of how similar two cell lines are in terms of 
\textit{local} spatial positionings.
Figure \ref{fig:orbits_correlations_all} 
demonstrates average correlations across all 72 orbits within each contact map.
Figure \ref{fig:orbits_correlations_all} clearly illustrates certain
orbits of Leukemic have higher correlation to each other than to the
normal MIT cell. In fact our statistical analysis shows that 
\textit {for orbits in $\mathbb{Q}$, 
intra-leukemic orbit correlations are significantly higher
than leukemic-normal orbit correlations}. This implies
there are significant differences between normal and
leukemic cells in terms of their local structure.
\begin{figure*}
    \centering
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{figures/contact_maps_correlations_cancer_vs_normal.png}
    \end{subfigure}
    \caption{
        \textbf{Pair-wise 
        average contact map orbit correlations
        for all contact maps:}
        ($\bar{\mathbf{m}}^{\scriptscriptstyle \scriptscriptstyle A,B}_{i,j} 
        \quad \forall i,j \in \{1 ... 23\} \quad \& \quad j \ge i$:
        average along the 
        red \textit{vertical} arrow in figure 
        \ref{graphlet_correlation_schema}).
        These values are calculated by averaging over 
        pairwise correlations of orbits of
        $\mathbb{Q}$ in a contact map.
}
    \label{fig:contact_map_correlations_all}
\end{figure*}

\begin{figure*}
    \centering
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{figures/orbits_correlations_cancer_vs_normal.png}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{figures/orbits_correlations_all.png}
    \end{subfigure}
    \caption{
        \textbf{Pair-wise average orbit correlations:}
        In figure \ref{fig:orbits_correlations_all}, each point
        in the graph is the result of averaging pair-wise
        orbit correlations over all contact maps
        ($\frac{1}{276}\sum_{i=0}^{23}\sum_{j=i}^{23}{
        m^{\scriptscriptstyle \scriptscriptstyle A,B}_{i,j, o}} \quad 
        \forall o \in \{0, 1, ..., 72\}$:
        average along the red \textit{horizontal} arrows in figure 
        \ref{graphlet_correlation_schema}).
        Counts for certain orbits are always zero in inter-chromosomal
        maps, leading to average value close to zero in 
        Figure \ref{fig:orbits_correlations_all}.
        In this figure, cancer cells data points are depicted
        in warm colors while normal cells are depicted in
        cold colors in increased contrast. As can be seen
        orbit distributions are more similar to each other
        for cancer cells.
    }
    \label{fig:orbits_correlations_all}
\end{figure*}

\subsection{Statistical Analysis}
In order to qauntify our results, we have conducted one-sided
t-test in order to test whether the average pair-wise
MIC values across orbits are significantly different from 
each other and also from 1. 
however, pairwise correlation of cancer cells for
orbits in $\mathbb{Q}$ were significantly higher than
correlation between cancer cells and normal cells.
The results
for this test showed that all values are significantly less than 1;
A summary of can be found in
table \ref{tab:t-test}.
We have also conducted t-tests for 
pair-wise orbit signature vector distances across contact maps.
Please refer to supplementary material for result of the 
full list of t-test results.
\begin{table}[H]
    \resizebox{.5\textwidth}{!}{
\begin{tabular}{c|ccccccc}
    & $MIC_{\scriptscriptstyle MIT-CALL4}$ & $MIC_{\scriptscriptstyle MIT-RL}$ & $MIC_{\scriptscriptstyle MIT-ALL}$ & $MIC_{\scriptscriptstyle ALL-CALL4}$ & $MIC_{\scriptscriptstyle CAll4-RL}$ & $MIC_{\scriptscriptstyle ALL-RL}$ & 1   \\
              \rule{0pt}{1pt} \\ \hline \rule{0pt}{15pt}
              $MIC_{\scriptscriptstyle MIT-CALL4}$ & -         & $\mathbf{<}$    & $\mathbf{<}$     & $\mathbf{<}$       & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle MIT-RL}$    & $\mathbf{>}$       &   -    & $\mathbf{<}$     & $\mathbf{<}$       & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle MIT-ALL}$   & $\mathbf{>}$       & $\mathbf{>}$    &    -    & $\mathbf{<}$       & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle ALL-CALL4}$ & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     &       -   & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle CAll4-RL}$  & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     & $\mathbf{>}$       &    -     & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle ALL-RL}$    & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     & $\mathbf{>}$       & $\mathbf{>}$      &   -    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              1         & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     & $\mathbf{>}$       & $\mathbf{>}$      & $\mathbf{>}$    &    -
\end{tabular}}
    \caption{
        Result of one-sided t-test 
        comparison of MIC results 
        with significance $\alpha = 0.01$
        for orbits of $\mathbb{Q}$.
        A $<$ sign means that the value in the corresponding
        row is significantly smaller that the value in 
        corresponding column; and a $>$ sign means the opposite.
        An $=$ sign would have meant the values are not 
        statistically different.
        As can be seen, all MIC values are significantly smaller that 1,
        meaning non on them are exactly the same; 
        however, all Cancer vs Cancer MIC values are statistically
        larger than Cancer vs Normal, implying that 
        spatial positioning of loci are more similar between cancer cells.
    }
    \label{tab:t-test}

\end{table}

>>>>

\section{results and discussions}
\subsection{Contact Map Orbit Vector Distance}
\begin{figure*}
    \centering
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{figures/orbit-distances_chr14.png}
    \end{subfigure}
    \caption{   
        \textbf{Pair-wise graphlet signature difference for
        all 276 contact maps:}
        Each point on a graph is the result of averaging all
        the distances across all loci of that contact map.
        ($\bar{\mathbf{d}}^{\scriptscriptstyle A,B}_{i,j} \quad
        \forall i,j \in \{1 ... 23\} \quad \& \quad j \ge i$).  
     }
    \label{fig:results_all}
\end{figure*}

@@

It is worth mentioning that interchromosomal 
thresholded contact maps 
represent
a bipartide graph with the loci from each chromosome 
on one side. Due to this
bipartide nature of the graphs in inter-chromosomal maps,
count of certain orbits is always 0, resulting in
a correlation values of 0 for them as well.
We ignored these values  when we calculated averages across orbits 
in figure \ref{fig:contact_maps_correlations_chr14} since they
would result in a bias towards zero in 
averages. You can see the bias in 
figure \ref{fig:orbits_correlations_all} where 
average correlations of orbits
$\mathbb{Q} = \{3, 9, 10{\text -}14, 20{\text -}34, 39{\text -}48, 51{\text -}72\}$ 
are close to zero. In fact all correlations
corresponding to these orbits are 0 except for the ones between the same 
chromosomes.

\subsection{MIC Comparison}
\begin{figure*}
    \centering
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{figures/contact_maps_correlations_cancer_vs_normal.png}
    \caption{}
    \label{fig:contact_map_correlations_all}
    \end{subfigure}

    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{figures/orbits_correlations_cancer_vs_normal.png}
    \caption{}
    \label{fig:orbits_correlations_normal_vs_cancer}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{figures/orbits_correlations_all.png}
    \caption{}
    \label{fig:orbits_correlations_all}
    \end{subfigure}
    \caption{
        \textbf{Pair-wise 
        average contact map orbit correlations
        for all contact maps:}
        ($\bar{\mathbf{m}}^{\scriptscriptstyle \scriptscriptstyle A,B}_{i,j} 
        \quad \forall i,j \in \{1 ... 23\} \quad \& \quad j \ge i$:
        average along the 
        red \textit{vertical} arrow in figure 
        \ref{graphlet_correlation_schema}).
        These values are calculated by averaging over 
        pairwise correlations of orbits of
        $\mathbb{Q}$ in a contact map.
        \textbf{Pair-wise average orbit correlations:}
        In figure \ref{fig:orbits_correlations_all}, each point
        in the graph is the result of averaging pair-wise
        orbit correlations over all contact maps
        ($\frac{1}{276}\sum_{i=0}^{23}\sum_{j=i}^{23}{
        m^{\scriptscriptstyle \scriptscriptstyle A,B}_{i,j, o}} \quad 
        \forall o \in \{0, 1, ..., 72\}$:
        average along the red \textit{horizontal} arrows in figure 
        \ref{graphlet_correlation_schema}).
        Counts for certain orbits are always zero in inter-chromosomal
        maps, leading to average value close to zero in 
        Figure \ref{fig:orbits_correlations_all}.
        In this figure, cancer cells data points are depicted
        in warm colors while normal cells are depicted in
        cold colors in increased contrast. As can be seen
        orbit distributions are more similar to each other
        for cancer cells.
    }
    \label{fig:orbits_correlations}
\end{figure*}

In addition to simple graphlet distance comparison,
we can compare graphlets by calculating
how often certain graphlets occur in them. By doing so
we measure the frequency distribution
of certain spatial structures in
each contact map. As a result of this approach
we can compare contact maps by calculating the 
correlation between their orbit distributions.
A higher correlation would mean higher similarity
in terms of spatial structure between the loci
involved.

This approach also allows for comparison of orbits
themsleves. By comparing average oribit correlations
across all contact maps of two cell lines, we can
have a measure of how similar certain orbits are
between cell lines. This can give us a 
\textit{global} measure
of how similar two cell lines are in terms of 
\textit{local} spatial positionings.

We caclulated pair-wise MIC values 
for each orbit in each of the 276
contact maps from MIT, ALL, RL, and CALL4 data separately. 
We found statistically significant difference between
normal vs cancer and cancer vs cancer correlations.
The results of staistical tests are shown in table
\ref{tab:statistical_results}. The difference
between the two cells are also illustrated in figure
\ref{fig:orbits_correlations}, which showes corrected
mean difference of MIC values. Figure 
\ref{fig:contact_map_correlations_all} shows the difference
between average pairwise
correlations. For most of the contact maps, correlation
between normal and cancer cells are smaller
than the correlation between cancer cells. This is corroborated
by the results of statical test which verify that such difference
is if fact signifcant.
Figures \ref{fig:orbits_correlations_normal_vs_cancer} and
\ref{fig:orbits_correlations_all} both show more details about
this difference in correlation. As can be observed, for obits
in $\mathbb{Q}$, normal correlations are smaller than cancer
correlations, while for the rest of the orbits, there is no
difference.

Figure \ref{fig:orbits_correlations} demonstrates that certain
orbits of Leukemic have higher correlation to each other 
than to the
normal MIT cell. In fact our statistical analysis shows that 
\textit {for orbits in $\mathbb{Q}$, 
intra-leukemic orbit correlations are significantly higher
than leukemic-normal orbit correlations}. This implies
there are significant differences between normal and
leukemic cells in terms of their local structure.

\subsection{Statistical Analysis}
In order to qauntify our results, we have conducted one-sided
t-test in order to test whether the average pair-wise
MIC values across orbits are significantly different from 
each other and also from 1. 
However, pairwise correlation of cancer cells for
orbits in $\mathbb{Q}$ were significantly higher than
correlation between cancer cells and normal cells.
The results
for this test showed that all values are significantly less than 1;
A summary of can be found in
table \ref{tab:t-test}.
We have also conducted t-tests for 
pair-wise orbit signature vector distances across contact maps.
Please refer to supplementary material for result of the 
full list of t-test results.
